---
title: "Open source software for reproducible transport data analysis: from zones to route networks"
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
library(sf)
```

```{r, eval=FALSE, echo=FALSE}
# Generate citations (requires Zotero)
library(rbbt)
bbt_write_bib(path = "references.bib", keys = bbt_detect_citations("README.Rmd"), overwrite = TRUE)
```


<!-- badges: start -->
[![.github/workflows/render-rmarkdown.yaml](https://github.com/Robinlovelace/odjitter/actions/workflows/render-rmarkdown.yaml/badge.svg)](https://github.com/Robinlovelace/odjitter/actions/workflows/render-rmarkdown.yaml)
<!-- badges: end -->


# Introduction

This repo was created to support at Open Data Manchester's event on open transport data, but it should be useful beyond that event, for anyone wanting to get, analyse and model transport data with open source software for transparent and evidence-based decision-making.

The amount of open data on transport systems can be overwhelming, especially when much of it is hard to download, let alone visualise and model and edit. In this talk I will talk about tools that can help with accessing open transport data to support you to generate new evidence and analysis in support of positive changes on vital travel networks in Chorlton and beyond. I will show how to download and work with data on road networks, road traffic casualties, and travel behaviour in R, a statistical programming language with outstanding visualisation, geographic analysis and statistical modelling capabilities. No need to 'live code' during the session: all the scripts to reproduce the outputs from the presentations will be provided in the open to support collaborative transport planning research.

# Set-up

To reproduce the code in this repo you will need to have R installed and, most likely, an IDE for R such as RStudio (recommended unless you already have a favourite coding tool that has good support for R such as VSCode).
You will also need to install some packages, e.g. by entering the following commands into the R console:

```{r, eval=FALSE}
pkgs = c(
  "pct",
  "stats19",
  "osmextract"
)
```

We will load the packages as we use them.

# Zone data from the PCT

```{r}
head(pct::pct_regions$region_name)
zones = pct::get_pct_zones("greater-manchester")
names(zones)[1:20]
names_to_plot = c("bicycle", "foot", "car_driver", "bus")
plot(zones[names_to_plot] )
```

# Desire line data from the PCT


# Crash data from stats19


# Transport infrastructure data from osmextra


# Scenarios of change

# Preparing data for A/B Street



# References

 
 
