---
title: "Open source software for reproducible transport data analysis: from zones to route networks"
output: github_document
bibliography: references.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
library(sf)
```

```{r, eval=FALSE, echo=FALSE}
# Generate citations (requires Zotero)
library(rbbt)
bbt_write_bib(path = "references.bib", keys = bbt_detect_citations("README.Rmd"), overwrite = TRUE)
```


<!-- badges: start -->
[![.github/workflows/render-rmarkdown.yaml](https://github.com/Robinlovelace/odjitter/actions/workflows/render-rmarkdown.yaml/badge.svg)](https://github.com/Robinlovelace/odjitter/actions/workflows/render-rmarkdown.yaml)
<!-- badges: end -->


# Introduction

This repo was created to support at Open Data Manchester's event on open transport data, but it should be useful beyond that event, for anyone wanting to get, analyse and model transport data with open source software for transparent and evidence-based decision-making.

The amount of open data on transport systems can be overwhelming, especially when much of it is hard to download, let alone visualise and model and edit. In this talk I will talk about tools that can help with accessing open transport data to support you to generate new evidence and analysis in support of positive changes on vital travel networks in Chorlton and beyond. I will show how to download and work with data on road networks, road traffic casualties, and travel behaviour in R, a statistical programming language with outstanding visualisation, geographic analysis and statistical modelling capabilities. No need to 'live code' during the session: all the scripts to reproduce the outputs from the presentations will be provided in the open to support collaborative transport planning research.

# Set-up

To reproduce the code in this repo you will need to have R installed and, most likely, an IDE for R such as RStudio (recommended unless you already have a favourite coding tool that has good support for R such as VSCode).

If you're new to R, it may be worth reading up on introductory material such as the free and open source resource *Reproducible Road Safety with R* [@lovelace_reproducible_2020] tutorial.
See [Section 1.5](https://itsleeds.github.io/rrsrr/introduction.html#installing-r-and-rstudio) of that tutorial to install R/RStudio and [Section 3](https://itsleeds.github.io/rrsrr/rstudio.html) on getting started with the powerful RStudio editor.
A strength of R is the number of high quality and open access [tutorials](https://education.rstudio.com/learn/beginner/), [books](https://education.rstudio.com/learn/beginner/) and videos to get started.

If you have R installed, you should be able to run all the code in this example and reproduce the results.

The first step is to install some packages, by entering the following commands into the R console:

```{r, eval=FALSE}
pkgs = c(
  "pct",
  "stats19",
  "osmextract",
  "tmap",
  "stplanr",
  "od"
)
```

You can install these packages as follows:

```{r, eval=FALSE}
install.packages(pkgs)
```

You can load these packages one-by-one with `library(pct)`, or all at once as follows:

```{r}
lapply(pkgs, library, character.only = TRUE)[length(pkgs)]
```

One final line of code to set-up the environment is to switch `tmap` into 'view' mode to create interactive maps:

```{r}
tmap_mode("view")
```



# Defining the study area

The first stage in many projects involving geographic data is defining the study area.
This is not always a straightforward or objective process.
In this case, the aim is to demonstrate how open data can be downloaded and visualised with a focus on Chorlton and with a view to getting the data into the transport simulation software A/B Street.

We will therefore select an area containing Chorlton and enough of the surrounding area to enable modelling of trips to key destinations.
As a starting point, we will use a 2 km buffer around the straight line between Chorlton and Manchester city centre to capture movement along this transport corridor:

```{r}
chorlton_point = tmaptools::geocode_OSM("chorlton, manchester")
manchester_point = tmaptools::geocode_OSM("manchester")
c_m_coordiantes = rbind(chorlton_point$coords, manchester_point$coords)
c_m_od = od::points_to_od(p = c_m_coordiantes, interzone_only = TRUE)
c_m_desire_line = od::odc_to_sf(c_m_od[-(1:2)])[1, ]
c_m_buffer = stplanr::geo_buffer(c_m_desire_line, dist = 2000)
qtm(c_m_buffer)
```


# Zone data from the PCT

```{r}
head(pct::pct_regions$region_name)
zones = pct::get_pct_zones("greater-manchester")
names(zones)[1:20]
names_to_plot = c("bicycle", "foot", "car_driver", "bus")
plot(zones[names_to_plot] )
```

# Desire line data from the PCT


# Crash data from stats19


# Transport infrastructure data from osmextra


# Scenarios of change

# Preparing data for A/B Street



# References

 
 
